---
title: "Correlations to Prior Periods"
output: html_document
date: "2023-08-17"
---

This script will attempt to ty to find prior period correlations with a specific date range set by the user.

```{r setup, include=FALSE}
require("quantmod");require("pbapply");require("data.table");require('highcharter');require("lubridate")
```

## SPX Daily Prices

Get stock OHLCV data and calculate returns.
```{r spyData}
# assign stock/etf/index
ticker = "^GSPC"
# stock prices
STK <- getSymbols(ticker,from="1900-01-01",auto.assign = FALSE)
# get returns
RETS <- na.omit(ROC(Cl(STK)))
# cumalative returns
cRETS <- cumsum(RETS)

```

## Test Range

This is the date range you want to test.
```{r assignDateRange}
# date range: "2020-01-01/2023-08-17"
date_range = paste0("2020-01-01/",Sys.Date())
# subset test range
test_range = cRETS[date_range]
# assign number of rows for the step size
tot_rows = nrow(test_range)
```


## Run Correlation Test
Steps through each day and returns correlations
```{r corTest}

# assign last obs
MAX = nrow(cRETS)-tot_rows
# step through the data by 1 day to calculate correlation
cor_ls = pblapply(as.list(0:MAX), function(ii){
  # we will step by row num
  tmp <- cRETS[(ii+1):(tot_rows+ii),]
  # run correlation test
  cor_res <- as.numeric(cor(tmp,test_range))
  # extract date range
  start_date <- index(tmp)[1]
  end_date <- index(tmp)[nrow(tmp)]
  # return data frame
  data.frame(start_date=start_date,
             end_date=end_date,
             cor = round(cor_res,5)
             )
})
# row bind all results
cor_res = as.data.frame(rbindlist(cor_ls,use.names = TRUE,fill = TRUE))
```

## Results

Depending on the date range you set, you may need to exclude dates "nearby". Since I used multi-year, I will remove `end_date` that are the same as the current year.
```{r corRes}
# remove this year's data
cor_res_sub = subset(cor_res,as.Date(cor_res$end_date) < as.Date("2023-01-01"))
# sort by correlation (highest to lowest)
cor_res_sub = cor_res_sub[order(cor_res_sub$cor,decreasing = TRUE),]
# plot the highest correlation with the test_range -- adding a year out 
high_cor <- cRETS[paste0(cor_res_sub$start_date[1],"/",as.Date(cor_res_sub$end_date[1])+lubridate::years(1))]
# reclass newer XTS
plotCors = merge(high_cor,xts(coredata(test_range),order.by = index(high_cor)[1:nrow(test_range)]))
colnames(plotCors) = c("highest_cor","current")

highcharter::highchart(type = "stock") %>%
  hc_title(text = "Correlation - Current Vs Prior") %>%
  hc_yAxis_multiples(list(title = list(text = "Return"), opposite = FALSE),
                       list(showLastLabel = FALSE, opposite = TRUE, title = list(text = "Current"))) %>%
  hc_add_series(plotCors$highest_cor, type="line",yAxis=0) %>% 
  hc_add_series(plotCors$current, type="line",yAxis=1) %>%
  hc_tooltip(pointFormat = '{point.x:%Y-%m-%d}
                            {point.y: .4f}%')

```

