---
title: "Volatility Trader"
output: html_document
date: "2023-08-01"
---

<https://cran.r-project.org/web/packages/volatilityTrader/volatilityTrader.pdf>
<https://cran.r-project.org/web/packages/bearishTrader/bearishTrader.pdf>
<https://cran.r-project.org/web/packages/bullishTrader/bullishTrader.pdf>

```{r setup, include=FALSE, echo=FALSE}
# load packages
require('volatilityTrader');require("dplyr")
```

## Volatility Trader Package Wrappers

1. `aStraddlePnL` : Calculates per share Profit and Loss (PnL) at expiration for Straddle Option Strategy and draws its Bar Plot displaying PnL in the Plots tab.

```{r  straddlePnL, echo=FALSE}
# QQQ options
# variable assignment
underlying_prc = 382.79
call_premium = 6.65
put_premium = 4.92
strike_prc = 382
lower_lmt = 0.95
upper_lmt = 1.05
# calculate PnL at Expiration
stradPnl <- aStraddlePnL(ST=underlying_prc,
             X = strike_prc,
             C = call_premium,
             P = put_premium,
             hl=lower_lmt,
             hu=upper_lmt
             )
# plot
print(stradPnl)
# get levels 
print(as.data.frame(stradPnl$data))
```

2. `strapPnL` = Calculates per share Profit and Loss (PnL) at expiration for Strap Option Strategy and draws its Bar Plot displaying PnL in the Plots tab.

This is a volatility strategy consisting of a long position in two ATM (at the money) calls , and a long position in an ATM (at the money) put option with a strike price X. This is a net debit trade.
The trader or investor has bullish outlook (Kakushadze & Serur, 2018).
```{r strapPlt, echo=FALSE}
# TSLA options
# variable assignment
underlying_prc = 261.07
strk = 260
call_premium1 = 14.40
call_premium2 = 14.40
put_premium = 12.00
lower_lmt = 0.70
upper_lmt = 1.30

strapPlt <- strapPnL(ST=underlying_prc,
                     X = strk,
                     C1 = call_premium1,
                     C2 = call_premium2,
                     P = put_premium,
                     hl = lower_lmt,
                     hu = upper_lmt
)

# plot
print(strapPlt)
# get levels 
print(as.data.frame(strapPlt$data))

```
